\LARGE \bf Attack Graph Generation for Micro-service Architecture

### 2018-08-02 paper_own_docker/main.tex
Proposed method section is complete. two images are added and Evaluation section is begun.

In this section, we first have a look at docker and attack graph terminology. We then present an overview of our proposed system with its components and at the end describe each of these components in more detail.

In Docker, distinction is being made between the terms image, container and service. Image is an executable package that includes everything needed to run an application,  container is a runtime instance of an image and service represents a container in production. A service only runs one image, but it codifies the way that image runs—what ports it should use, how many replicas of the container should run so the service has the capacity it needs, and so on. In our work we treat these terms equally, since we are doing a statical attack graph analysis.

In this work we model an attack graph as a sequence of atomic attacks. Each atomic attack represents a transition from a component(with its privilege) to either the same component with a higher privilege or a completely new component with a new privilege. A goal of an attacker would be to perform multiple atomic attacks to get to the desired goal container. For example, let us suppose that an attacker wants to have access to a database of a certain website. In order to reach the database, he has to pass other containers between them/ He does that by finding a vulnerability to exploit and give access to the container's neighbors.After a successful exploitation of a chain of intermediate containers, he finally has access to the database and its functions. Eventhough attack graphs model the attacker scenario, they are of crucial importance in computer security. A system administrator would be interested to have an overview of the attack paths that an attacker could exploit, in order to harden the security of a given enterprise system.

@@

Atomic attacks are the fundamental building blocks in our attack graph. A single atomic attack represents a successful vulnerability exploitation. A consecutive sequence of atomic attacks represents a path in an attack graph and multiple attack paths constitute the attack graph. For an atomic attack to be executed, two constraints are imposed. First, the containers have to be connected to each other, so that physical access can be ensured. Second, the attacked container should contain some vulnerability that the attacker can exploit and gain a privilege level. We model the privileges into a hierarchy. The privileges in ascending order are: None, VOS(User), VOS(Admin), OS(User) and OS(Admin). VOS means that the privilege is connected to a virtual machine, while OS means that the host machine has been infected. Since VOS are on some hosts, they have lower level of hierarchy.

Furthermore, in order for an atomic attack to be performed, a certain precondition has to be met. Preconditions are privilege levels that are required so that a vulnerability can be exploited. When an atomic attack is successfully executed, a postcondition is obtained. Postconditions are privileges acquired as a result of a successful attack. Both the Pre- and Postconditions are transformed from pre- and postcondition rules which are collected by experts.

Nodes in our attack graph model are composed as a combination of (docker image and privilege level), while edges are a connection between node pairs with a vulnerability as a descriptor. Once an attacker exploits a given vulnerability, he gains the privileges for the new container and an edge is added to the attack graph.

Our attack graph generator is composed of three main components: Topology Parser, Vulnerability Parser and Attack Graph Parser(Fig. \ref{AttackGraphSystem}). The Topology Parser reads the underlying topology of the system and converts it to a more machine readable format, the Vulnerability Parser generates the vulnerabilities for each of the images and the Attack Graph Parser from the topology and vulnerabilities files generates the attack graph.

<INSERT IMAGE PRIVILEGES><REF new paper>

<INSERT IMAGE EXAMPLE>

<INSERT IMAGE ATTACK GRAPH>

In the following subsections, we first look into the required technical system specifications, we then describe each of the parsers in more detail and at the end examine the characteristics of the Breath-First Search used by the Attack Graph Parser.

@@

\subsubsection{Topology Parser}
The topology of Docker containers can be described in either runtime, or by using Docker Compose. In our case, since we are doing static attack graph analysis, we use Docker Compose. Docker Compose provides us with a docker-compose.yml which is used for extraction of the topology of the system. Even docker-compose.yml are different, based on the version. For example older versions use the deprecated keyword link, while newer ones use exclusively networks. In this work, we use the keyword networks as a statement that a connection between two containers exists.

However, in the majority of cases, in order for an application to be useful, it has to communicate with the outside world. The containers that have published ports are connected to the outside.

Some containers have privileged access. That means that an attacker with access to these containers, has also access to the docker daemon. This can be done by us

\subsubsection{Vulnerability Parser}
In the preprocessing step, we use Clair to generate the vulnerabilities of a given container. Clair provides CVE-ID, description and attack vector for each vulnerability. Attack vector is an entity that describes which conditions and effects are connected to this vulnerability. The fields in the attack vector as described by the National Vulnerability Database(NVD) are: Access Vector(Local, Adjacent Network and Network), Access Complexity(Low, Medium, High), Authentication(None, Single, Multiple), Confidentiality Impact(None, Partial, Complete), Integrity Impact(None, Partial, Complete) and Availability Impact(None Partial Complete). However, Clair does not provide with an easy to use interface to analyze a docker image. As a result, we use Clairctl(Clair wrapper) in order to analyze a complete docker image.

We here first preprocess the vulnerabilities and convert them into sets of pre- and postconditions. In order to do this, we match the attack vectors acquired earlier from the vulnerability database and keywords of the descriptions of each vulnerability to generate attack rules. When a subset of attack vector fields and description keywords match a given rule, we use the pre- or postcondition of that rule. If more than one rule match, we take the one with the highest privilege level for the preconditions and the lowest privilege level for the postconditions. If no rule matches we take None as a precondition and ADMIN(OS) as a postcondition.This results in a list of container vulnerabilities with their preconditions and postconditions.
\subsubsection{Breadth-First Search}

After the preprocessing step is done, the vulnerabilities are parsed and their pre- and postconditions are extracted. Together with the topology, they are feed into the Breadth-First Search algorithm(BFS).
Breadth-first search is a popular search algorithm that traverses a graph by looking first at the neighbors of a given node, before going deeper in the graph. Pseudocode of our modified Breath First Search is given on (Figure.)

@@

\caption{Breadth-first search algorithm for generating an attack graph.}
\end{algorithm}

The algorithm requires the topology and a dictionary of the exploitable vulnerabilities as an input and the output is made up of the nodes and the edges that make the attack graph. 
The algorithm first initializes the nodes, edges, queue and the passed nodes. Afterwards it generates the nodes which are a combination of the image name and the privilege level.
Then into a while loop we iterate through every node, check its neighbors and add the edges. If the neighbor was not passed, then we add it to the queue. The algorithm terminates when the queue is empty.

Breath First Search is characterized by the following properties:

\begin{itemize}
 \item Completeness: Breadth First Search is complete i.e. if there is a solution, breadth-first search will find it regardless of the kind of graph.
 \item Termination: This follows from its monotonicity property. Each edge is traversed only once.
 \item Time Complexity: is $O(|N| + |E|)$ where $|N|$ is the number of nodes and $|E|$ is the number of edges in the attack graph.
\end{itemize}
\section{EVALUATION}

In this section, we will conduct few experiments in order to test the scalability of our system with different number of containers and links between them.

Throughout the experiments, the biggest time bottleneck is the preprocessing step, and the graph drawing step. However these steps are with linear complexity because the container files are analyzied only once. The attack graph generation less time than the preprocessing step.

The following experiments were perfomed. We used the Samba(reference) and Phpmailer(reference) examples which were taken from . Where we artificially made clique of 1, 5, 20, 50, 100, 500 and 1000 containers to test the scalability of the system. The Phpmailer container has 181 vulnerabilities, while the Samba container has 367 vulnerabilities detected by Clair.

On the table(Table I) we can see the results of our experiments. In each of the experiments the number of Phpcontainer is constant, while the number of Samba containers is increasing in a clique fashion. There are also two artifitial containers("outside" that represents the outside world from where the attacker can attack and the "docker host", i.e. the docker daemon where the containers are present). Therefore the number of nodes in the topology graph is the sum of: "outside", "docker host", number of Phpmailer containers and number of Samba containers. The number of edges of the topology graph is: 1 edge("outside"-"Phpmailer"), n edges("docker host" to all of the containers) and clique of the Phpmailer and samba containers n*(n+1)/2. For example\_5, the number of edges in the topology graph would be 32: 1 outside edge, 6 docker host edges(n=6, 1 Phpmailer and 5 Sambas) and 25 clicque edges(5*6/2=15).

We also performed testing on a real example- atsea sample shop app(reference). The system is composed of the containers app, database, payment\_gateway and reverse\_proxy.

@@

	\caption{List of randomly selected examples that were analyzed with our attack graph generation system.}

@@

		No. of nodes in topology & 4 & 8 & 23 & 53 & 103 & 503 & n\\

	    No. of edges in topology & 6 & 28 & 253 & 1378 & 5253 & 126253 & n \\

		No. nodes in attack graph & 5 & 13 & 43 & 103 & 203 & 1003 & n \\

		No. edges in attack graph & 8 & 68 & 863 & 5153 & 20303 & 501503 & n \\

		Topology parsing time & 0.0082 & 0.0094 & 0.02879 & 0.0563 & 0.1241 & 0.7184 & n \\

		Vulnerabilities preprocessing time & 0.2551 & 0.2840 & 0.5377 & 0.9128 & 1.6648 & 6.9961 & n \\

		Breadth-First Search time & 0.0019 & 0.0209 & 0.2763 & 1.6524 & 6.5527 & 165.3634 & n \\

		Total time & 0.2654 & 0.3144 & 0.8429 & 2.6216 & 8.3417 & 173.0781 & n \\ 
		\hline
	\end{tabular}
\end{center}

>>>>

In this section we first have a look at Docker and attack graph terminology. We then present an overview of our proposed system with its components, provide a small example and at the end describe each of the components in more detail.

In Docker, distinction is being made between the terms image, container and service. Image is an executable package that includes everything needed to run an application,  container is a runtime instance of an image and service represents a container in production. A service only runs one image, but it codifies the way that image runs—what ports it should use, how many replicas of the container should run so the service has the capacity it needs, and so on. In our work we treat these terms equally, since we are doing a static and not runtime attack graph analysis.

In this work we model an attack graph as a sequence of atomic attacks. Each atomic attack represents a transition from a component(with its privilege) to either the same component with a higher privilege or to a neighbor component with a new privilege. A goal of an attacker would be to perform multiple atomic attacks to get to the desired goal container. For example, let us suppose that an attacker wants to have access to a database of a certain website. In order to reach the database, he has to pass other containers between him and the database. He does that by finding a vulnerability to exploit and give access to the container's neighbors. After a successful exploitation of a chain of intermediate containers, he finally has access to the database and its functions. Even though attack graphs model the attacker scenario from an attackers perspective, they are of crucial importance in computer security. For example, a system administrator would be interested to have an overview of the attack paths that an attacker could exploit, in order to harden the security of a given enterprise system.

@@

Atomic attacks are fundamental building blocks in our attack graph. A single atomic attack represents a successful vulnerability exploitation. A consecutive sequence of atomic attacks represents a path in an attack graph and multiple attack paths constitute an attack graph. For an atomic attack to be executed, two constraints are imposed. First, the containers have to be connected between each other, so that physical access can be ensured. Second, the attacked container should contain some vulnerability that the attacker can exploit and gain a privilege level. We model the privileges into a hierarchy. The privileges in ascending order are: None, VOS(User), VOS(Admin), OS(User) and OS(Admin). VOS means that the privilege is connected to a virtual machine, while OS means that the host machine has been infected. Since VOS are on some hosts, they are in the lower level of hierarchy.

Furthermore, in order for an atomic attack to be performed, a certain precondition has to be met. Preconditions are privilege levels that are required so that a vulnerability can be exploited. When an atomic attack is successfully executed, a postcondition is obtained. Postconditions are privileges acquired as a result of a successful attack. Both the Pre- and Postconditions are transformed from pre- and postcondition rules. We use in our attack graph generator the pre- and postconditions from (reference) which are manually selected by experts and evaluated.

Nodes in our attack graph model are represented as a combination of a docker image and its respective privilege level, while edges are a connection between node pairs with the vulnerability that is being exploited as a descriptor. Once an attacker exploits a given vulnerability, he gains the privilege of the new container and an edge is added to the attack graph.

In order to show how the attack graph generation works in practice, we present a small example. The example is taken from the Netflix OSS Github repository. Displayed on figure \ref{TopologyGraph} is the topology of the example. The topology consists of "Outside" node, "Docker daemon" and the a subset of the containers. On figure \ref{AttackGraph}, we can see a part of the resulting attack graph. An example path than an attacker to take would be to first attack the Zuul container and gain USER privilege. Then with this USER privilege it can exploit another vulnerability to gain ADMIN privilege. Once the ADMIN privilege has been obtained on Zuul, the attacker can attack the Eureka container and gain ADMIN privilege. It is important to note that this is not the only path that the attacker can take in order to have ADMIN privileges on Eureka. Another path would be to use another Zuul vulnerability to gain directly ADMIN privileges and then attack the Eureka container.

Our attack graph generator is composed of three main components: Topology Parser, Vulnerability Parser and Attack Graph Parser(Fig. \ref{AttackGraphSystem}). The Topology Parser reads the underlying topology of the system and converts it into a more machine readable format, the Vulnerability Parser generates the vulnerabilities for each of the images and the Attack Graph Parser generates the attack graph from the topology and vulnerabilities files.

\begin{figure}
	\includegraphics[]{Topology_graph}
	\caption{Example topology graph. The topology graph is a subset of a real topology graph from the Netflix OSS example. Each node denote container(plus Docker Deamon and Outside) and each edge denotes a connection between two containers.}
	\label{TopologyGraph}
\end{figure}

\begin{figure}
	\includegraphics[]{Attack_graph}
	\caption{Example resulting attack graph. This attack graph is a subset of a real attack graph from the Netflix OSS example. Nodes correspond to a pair of container plus privilege, while edges are atomic attacks.}
	\label{AttackGraph}
\end{figure}

In the following subsections, we first have a look into the system requirements, then describe each of the parsers in more detail and finally examine the characteristics of the Breath-First Search graph transversal algorithm.

@@

\subsubsection{Topology Parser}
The topology of Docker containers can be described at either runtime, or statically by using Docker Compose. In our case, since we are doing static attack graph analysis, we use Docker Compose as our main tool in the beginning. Docker Compose provides us with a docker-compose.yml file which is used for extraction of the topology of the system. However different version of docker-compose.yml, use different syntax. For example older versions use the deprecated keyword "link", while newer ones use exclusively "networks", to denote a connection between two containers. In this work, we use the keyword "networks" as an indicator that a connection between two containers exists.

However, in the majority of cases, in order for an application to be useful, it has to communicate with the outside world. This is usually done using publishing ports. This is the case in both computer networks, as well as in microservice architectures.

Another thing that we take into account is the privileged access. Some containers require certain privileges over the docker daemon in order to function properly. In docker this is usually done either by mounting the docker socket or specifying the keyword "privileged" in the docker-compose.yml file . An attacker with access to these containers, has also access to the docker daemon. And once he has access to the docker daemon, he has potentially access to the whole microservice system, since every container is controlled and hosted by the daemon.

@@

\caption{Breadth-first search algorithm for generating an attack graph.}
\label{BFSalgorithm}
\end{algorithm}

			No. of nodes in topology & 4 & 8 & 23 & 53 & 103 & 503 & 1003\\

			No. of edges in topology & 6 & 28 & 253 & 1378 & 5253 & 126253 & 502503 \\

			No. nodes in attack graph & 5 & 13 & 43 & 103 & 203 & 1003 & 2003 \\

			No. edges in attack graph & 8 & 68 & 863 & 5153 & 20303 & 501503 & 2003003 \\

			Topology parsing time & 0.0082 & 0.0094 & 0.02879 & 0.0563 & 0.1241 & 0.7184 & 2.3664 \\

			Vulnerabilities preprocessing time & 0.2551 & 0.2840 & 0.5377 & 0.9128 & 1.6648 & 6.9961 & 15.0639 \\

			Breadth-First Search time & 0.0019 & 0.0209 & 0.2763 & 1.6524 & 6.5527 & 165.3634 & 767.5539 \\

			Total time & 0.2654 & 0.3144 & 0.8429 & 2.6216 & 8.3417 & 173.0781 & 784.9843 \\ 
			\hline
		\end{tabular}
	\end{center}

	\label{table_scalability}
\end{table*}[t]

\subsubsection{Vulnerability Parser}
In the preprocessing step, we use Clair to generate the vulnerabilities of a given container. Clair is a tool that inspects a docker image, and generate its vulnerabilities by providing CVE-ID, description and attack vector for each vulnerability. Attack vector is an entity that describes which conditions and effects are connected to this vulnerability. The fields in the attack vector as described by the National Vulnerability Database(NVD) are: Access Vector(Local, Adjacent Network and Network), Access Complexity(Low, Medium, High), Authentication(None, Single, Multiple), Confidentiality Impact(None, Partial, Complete), Integrity Impact(None, Partial, Complete) and Availability Impact(None Partial Complete). Although it is very useful, Clair does not provide with an easy to use interface to analyze a docker image. As a result, we use Clairctl(Clair wrapper) in order to analyze a complete docker image.

We here first preprocess the vulnerabilities and convert them into sets of pre- and postconditions. In order to do this, we match the attack vectors acquired earlier from the vulnerability database and keywords of the descriptions of each vulnerability to generate attack rules. When a subset of attack vector fields and description keywords match a given rule, we use the pre- or postcondition of that rule. If more than one rule match, we take the one with the highest privilege level for the preconditions and the lowest privilege level for the postconditions. If no rule matches we take None as a precondition and ADMIN(OS) as a postcondition.This results in a list of container vulnerabilities with their preconditions and postconditions.

\subsubsection{Breadth-First Search}

After the preprocessing step is done, the vulnerabilities are parsed and their pre- and postconditions are extracted. Together with the topology, they are feed into the Breadth-First Search algorithm(BFS).
Breadth-first search is a popular search algorithm that traverses a graph by looking first at the neighbors of a given node, before going deeper in the graph. Pseudocode of our modified Breath First Search is given in Algorithm \ref{BFSalgorithm}.

The algorithm requires the topology and a dictionary of the exploitable vulnerabilities as an input and the output is made up of the nodes and the edges that make the attack graph. 
The algorithm first initializes the nodes, edges, queue and the passed nodes. Afterwards it generates the nodes which are a combination of the image name and the privilege level.
Then into a while loop we iterate through every node, check its neighbors and add the edges. If the neighbor was not passed, then we add it to the queue. The algorithm terminates when the queue is empty. It is characterized by the following properties:

\begin{itemize}
 \item Completeness: Breadth First Search is complete i.e. if there is a solution, breadth-first search will find it regardless of the kind of graph.
 \item Termination: This follows from its monotonicity property. Each edge is traversed only once.
 \item Time Complexity: is $O(|N| + |E|)$ where $|N|$ is the number of nodes and $|E|$ is the number of edges in the attack graph.
\end{itemize}

\section{EVALUATION}

An attack graph system needs to be scalable because in a real world scenario, there are many containers interconnected to each other. (some references and description with statistics). In this section, we first have a look at how others evaluate their proposed systems. We will then conduct few experiments in order to test the scalability of our system with different number of containers and links between them in both synthetically-made and real networks.

Evaluation of a given system is a usually required part, when evaluating a given program. Extensive scalability study is not available. Characteristics that vary are the number of nodes, their interconnectedness and the amount of vulnerabilities per container. All of these components contribute to the execution time. Even though the notion of attack graph vary, we hope to reach a comprehensive comparison. In our case we treat every container as a computer, and treat any physical connection between two computers as a connection between two containers.

@@

	\caption{List of randomly selected examples that were analyzed with our attack graph generation system.}
	\label{table_technologies}

@@

For each example describe their setup and their results.

Seyner, the attack graph has 5948 nodes and 68364 edges. The time needed for NuSMV to execute this configuration is 2 hours, but the model checking part took 4 minutes. They claim that the performance bottleneck is inside the graph generation procedure.(reference) Therefore model checkers are not the most suitable graph representation.

Ingols tested on a network of 250 hosts. He also tested on a simulated network of 50000 hosts in under 4 minutes. (reference) However he does this evaluation on the Multiple Prerequisite graph, which is different from ours.

Amman does not provide with scalability methods.

Evaluation pre/postconditions where achieved, but they provide no evaluation for attack graph generation.

In our scalability experiments we use Samba(reference) and Phpmailer(reference) containers which were taken from their respective Github repositories. We then extended this example and artificially made cliques of 5, 20, 50, 100, 500 and 1000 Samba containers to test the scalability of the system. The Phpmailer container has 181 vulnerabilities, while the Samba container has 367 vulnerabilities detected by Clair.

On Table \ref{table_scalability} we can see the results of our experiments. In each of the experiments the number of Phpmailer containers stays constant, while the number of Samba containers is increasing. This increase is done in a clique fashion, a node of each additional container is connected to every existing container. In addition, there are also two additional artificial containers("outside" that represents the environment from where the attacker can attack and the "docker host", i.e. the docker daemon where the containers are present). Therefore the number of nodes in the topology graph is the sum of: "outside", "docker host", number of Phpmailer containers and number of Samba containers. The number of edges of the topology graph is a a combination of: 1 edge("outside"-"Phpmailer"), n edges("docker host" to all of the containers) and clique of the Phpmailer and samba containers n*(n+1)/2. For example\_5, the number of containers would be 8(1 Phpmailer, 1 outside, 1 docker host and 5 Samba containers) the number of edges in the topology graph would be 32: 1 outside edge, 6 docker host edges(n=6, 1 Phpmailer and 5 Sambas) and 25 clique edges(5*6/2=15).

Throughout the experiments, for the less connected ones, the biggest time bottleneck is the preprocessing step. However these steps are with the number of containers because the container files are analyzed only once. The attack graph generation less time than the preprocessing step. Starting from example\_500, we can notice sharp increase in the number of the BDF time to 165 seconds. For the previous example with example\_100, needed 6.5 seconds.

We also performed tests on some real examples as described on table. These examples are different from the synthetic ones presented above, since they contain less containers. Also they are in a more linear fashion. For example, in order for an attacker to reach the database, he needs to gain suitable privilege levels of multiple intermediate containers.

 We tested our generator on few other randomly selected examples from Github: NetflixOSS, Atsea Sample Shop App, and JavaEE demo. These systems are composed of different kind of technologies, different number of containers and vulnerabilities. Short summary of them is given on the table \ref{table_technologies}. We want to show that we can generalize our approach for different types of technologies.

Characteristics of the Attack Graph Generator.
 - stopping criteria (When all edges are traversed)
 - BFS is guaranteed to traverse all of the edges 
 - monothonicity no edge is visited more than once.
 - all of the vulnerabilities per edge are checked
 - only the ones that fulfill a certain precondition are added. If precondition is not fulfilled, then the edge is not added.
